### 2022-01-05 Script/2 Analyses.Rmd
add results

#ML Analysis

used libraries
```{r}
library(glmnet)
```

load in data
```{r}
#load in data minus csv-appended column
TextFeatures <- read.csv2(file = "../Data/6 TextFeatures.csv", sep = ",")[,-1]

#convert to numeric
TextFeatures <- sapply(TextFeatures, as.numeric) 
```

split data into train and test
```{r}
#seed for reproducability
set.seed(6164900)

#proportion of complete set dedicated to training set and split into test and training data
prop.train <- 0.6 
splits <- sample(1:nrow(TextFeatures), size = floor(nrow(TextFeatures)*prop.train))

train <- TextFeatures[splits, ]
test <- TextFeatures[-splits, ]

K <- 10 #number of folds

```

evaluation metrics and other functions used
```{r}
source("Functions.R")
```

###Multilinear Regression. 
Does not need CV in this case, no parameters to be tuned.
```{r}
fit_lm <- lm(Rating ~ ., data = as.data.frame(train)) #train model
pred_lm <- predict(fit_lm, newdata = as.data.frame(test)) #test model
metrics_lm <- metrics(test[,1], pred_lm) #obtain metrics

CIs <- confint(fit_lm, level = 0.95) #save CI
unimp <- rownames(CIs[CIs[,1] < 0 & CIs[,2] >0, ]) #unimportant predictors
imp_lm <- rownames(CIs)[!rownames(CIs) %in% unimp][-1] #important predictors
```

###lasso
```{r}
set.seed(6164900)
Lassocv <- cv.glmnet(x = train[, 2:ncol(train)], y = train[,1], alpha = 1, family = 'gaussian', standardize = F,
                    nfolds = K) #setting alpha to 1, runs Lasso

#evaluate cross validation
plot(Lassocv)

#train model
fit.Lasso <- glmnet(x = train[, 2:ncol(train)], y = train[,1], alpha = 1, family = 'gaussian', standardize = F,
                    lambda = 1.5)

#obtain important predictors
lasmat <- as.matrix(coef(fit.Lasso))
imp_Lasso <- names(lasmat[lasmat != 0,])[-1]

#evaluate model
pred_lasso <- predict(fit.Lasso, newx = test[,2:ncol(test)])
metrics_lasso <- metrics(test[,1], pred_lasso)

```

###Ridge
```{r}
K <- 10 #number of CVs

set.seed(6164900)
Ridgecv <- cv.glmnet(x = train[, 2:ncol(train)], y = train[,1], alpha = 0, family = 'gaussian', standardize = F,
                    nfolds = K) #setting alpha to 0, runs Ridge.

plot(Ridgecv) #lowest MSE is when lambda = its lowest value. also variance is not that big at that value

#train model
fit.Ridge <- glmnet(x = train[, 2:ncol(train)], y = train[,1], alpha = 0, family = 'gaussian', standardize = F,
                    lambda = Ridgecv$lambda.min)

#variable importance
ridgemat <- as.matrix(coef(fit.Ridge))
imp_Ridge <- names(ridgemat[abs(ridgemat) > 0.05,])[-1]

#evaluate model
pred_ridge <- predict(fit.Ridge, newx = test[,2:ncol(test)])
metrics_ridge <- metrics(test[,1], pred_ridge)

```

###Elastic Net
```{r}
#to cross-validate alpha
alpha <- seq(0.1, 0.9, by = 0.1)

#cross-validation of alpha
set.seed(6164900)
glmnetcvs <- list()
for (i in alpha) {
    glmnetcvs[[which(alpha == i)]] <- cv.glmnet(x = train[, 2:ncol(train)], y = train[,1], family = 'gaussian',
                                              standardize = F, nfolds = K, alpha = i)
}
names(glmnetcvs) <- sprintf("alpha%g", alpha)

#obtain lowest lambda and check variance of model at that lambda by looking at the plot
minlambda <- sapply(glmnetcvs, function(x){
  plot(x)
  x$lambda.min
})
min.lambda <- minlambda[minlambda == min(minlambda)] #alpha = 0.9, lambda = 0.0029

#train model
fit.EN <- glmnet(x = train[, 2:ncol(train)], y = train[,1], alpha = 0.9, family = 'gaussian', standardize = F,
                    lambda = min.lambda)

#variable importance
ENmat <- as.matrix(coef(fit.EN))
imp_EN <- names(ENmat[abs(ENmat) > 0.05,])[-1]

#evaluate model
pred_EN <- predict(fit.EN, newx = test[,2:ncol(test)])
metrics_EN <- metrics(test[,1], pred_EN)

```

###LPCR
preprocess for LPCR
```{r}
train <- cbind(train, ind = sample(rep(1:K, length.out = nrow(train)))) #indicator variable
traindf <- as.data.frame(train)
testdf <- as.data.frame(test)

```

Linear Principle Component regression algorithm.
```{r}
npc <- 1:4 #hyperparameters to be cross-validated

metrics_lpcr <- list() #save results of each condition in a list
for(j in npc){ #for every condition
  metric_cv <- matrix(NA, K, 3) #save results for a fold in matrix
  colnames(metric_cv) <- c("RMSE", "MAD", "R2") #give comprehensive names
  for(i in 1:K){ #for everly fold
    traincv <- traindf[traindf$ind != i,] #set training set
    validation <- traindf[traindf$ind == i,] #set testing set

   metric_cv[i,] <- LPCR(traincv, validation, npc = j, relevance = F) #run LPCR to obtain only metrics and save
  }
  metrics_lpcr[[j-(npc[1]-1)]] <- metric_cv #save in storage
}

names(metrics_lpcr) <- sprintf("NPC%d", npc) #give comprehensive names

```

LPCR obtain best condition
```{r}
eval.pcr <- evaluation(metrics_lpcr) #evaluate the list of conditions
eval.pcr$evaluation #list of metrics per condition
eval.pcr$RMSE #not much difference between 3 and 4
eval.pcr$lowest_RMSE #using 4 PC's gives the lowest RMSE, however it is not much better than when 3 PC's are used, so 3 is preferred. 
```

LPCR test dataset
```{r}
fit_lpcr <- LPCR(traindf, testdf, npc = 3, relevance = T) #set relevance to true returns variable importance
metrics_lpcr <- fit_lpcr$metrics
imp_lpcr <- fit_lpcr$imp.pcr
```

###obtain final results
```{r}
#importance_lst <- ls(pattern = "imp_")
#metrics_lst <- ls(pattern = "metrics_")
#remove all variables except the ones we still need
#rm(list=setdiff(ls(), c(importance_lst, metrics_lst)))

#obtain metrics
result_metrics <- do.call(rbind, mget(ls(pattern = "metrics_")))

#obtain variable importance
result_feature_importance <- table(do.call(c, mget(ls(pattern = "imp_"))))

results <- list(performance = result_metrics, feature_imp = result_feature_importance)

saveRDS(results, "../Results/Results.RData")

```

###delete environment and console if desired
```{r}
rm(list=ls())
cat("\014")
```